---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# dtplyr

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/mcskinner/dtplyr.svg?branch=master)](https://travis-ci.org/mcskinner/dtplyr)
<!-- badges: end -->

The [tidyverse](https://www.tidyverse.org) has made it far easier to perform data work in R.
But there are still times where performance is important, and for tabular data that's when
[data.table](https://github.com/Rdatatable/data.table) can really shine.

For some reason, however, there seems to be an either / or mentality when it comes to picking between `data.table` and `dplyr`.
I don't buy it. I'd rather have the best of both worlds.

That's why I made these handy adapters. Stick with your tidy data flow, but go ahead and cherry-pick some performance wins when you need them.

## Installation

**NOTE:** This is a lie right now. Dare to dream.

You can install the released version of dtplyr from [CRAN](https://CRAN.R-project.org) with:

```{r, eval = FALSE}
install.packages("dtplyr")
```

### Development version

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("mcskinner/dtplyr")
```

## Usage

Aggregation with `dt_agg()` uses the lean syntax you're used to from `data.table`:

```{r}
library(dtplyr)
dt_agg(iris, mean(Sepal.Length), Species)
```

You'll notice that aggregation skipped the filter. That's exposed by `dt_op()` if you need it back:

```{r}
dt_op(iris, Sepal.Width >= 3, mean(Sepal.Length), Species)
```

And why not leverage the high-performance merge routine as well?

```{r}
colors <- data.frame(Species = c('setosa', 'versicolor', 'virginica'), Color = c('baby blue', 'lavender', 'purple'))
head(dt_merge(iris, colors))
```
